---
title: "Predicting NBA Minutes played"
Author: "Robert Miller"
output: html_document
date: "2022-12-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width = 10, fig.height = 5)

devtools::install_github("abresler/nbastatR")
library(nbastatR)
library(tidymodels)
library(corrplot)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(pracma)
```
Data cleaning

```{r, eval=FALSE}
seq <- seq(22100001,22101320,8)
box_team<- box_scores(seq,box_score_types = c("Traditional", "Advanced"),result_types = "team", join_data = TRUE) 

get_seasons_rosters(seasons = 2021:2022, return_message = TRUE, nest_data = F)

save(box_team,file="box_team.rda")

box_player <-  box_scores(seq,box_score_types = c("Traditional", "Advanced"),result_types = "player", join_data = TRUE)
save(box_player,file = "box_player.rda")

```
```{r}

load(file = "box_team.rda")

load(file = "box_player.rda")








  
  
```


```{r}
#removes game id's, player id's, amd team id's
player <- select(box_player[[2]][[1]],-c(idGame,idTeam,idPlayer))




correlation <- cor(select_if(player,is.numeric))


corrplot(correlation,type = 'lower', diag = FALSE, method='color')

```

```{r}
ggplot(player,aes(minExact,pts))+ geom_point(aes(colour = cut(pctFG, c(0.0,0.25, 0.461,.667, 1.0))),size = 5) + facet_wrap(~groupStartPosition) 

```
Basketball is grouped into 3 groups (guards, forwards, centers) and 5 positions (point guard, shooting guard, small forward, power forward, and center). Player listed as NA do not have a starting position in the game and played off the bench. As we can see, the more minutes a player plays there is an upward trajectory in points scored. This makes sense, as its often the case where the player that score the most points will play the most minutes.  What about 3 point shots? They are quickly dominating the NBA and 3pt attempts are up league-wide. Is there a relationship between 3pt attempts and minutes played? what about 3pts attempted?  
```{r}

ggplot(player, aes(minExact,pctFG3)) + geom_point(aes(colour = cut(fg3a,c(-1,0,2, 4,6,8,10,15))),size = 5) + facet_wrap(~groupStartPosition)


```
As we can see both 3 point percentage and attempts is very dependent on minutes played. Centers shoot much less threes than guards and even forwards as a result of their often larger size prefer to shoot around the rim. We can also see that as minutes increase there is a decrease in 3 point efficiency. There are bands around 0%,50%, 100% as a result of players taking only a couple threes and hitting or missing all of them. In the 2021-2022 season the league wide average shooting percentage was 35.4% (source:basketballreference.com)


```{r}

ggplot(player, aes(minExact,netrtg)) + geom_point() + facet_wrap(~groupStartPosition) 



```

Net rating is defined as offensive rating vs defensive rating. Offensive rating is a measure of how may points are scored by the player per 100 possessions of the basketball and defensive rating is measure of how many points they give up on defense per 100 possessions. If more points are scored then given up, this will result in a positive net rating. The best players often average a high positive net rating, for example Nikola Jokic a 6'11 Serbian Center that plays for the Denver Nuggets averaged net ratings of 14.8 in 20-21 season and 10.4 in 21-22 season, earning him back to back most valuable player awards. High net ratings may be a good predictor of many minutes played, but astronomically high or low net ratings are a result of a small sample size which may be a good predictor for players playing few minutes. 